---
title: "*RBGL*: R interface to boost graph library"
author: 
- name: "L. Long"
- name: "VJ Carey"
- name: "R. Gentleman"
- name: "Emmanuel Taiwo"
  affiliation: "Vignette translation from Sweave to R Markdown / HTML"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{RBGL Overview}
  %\VignetteDepends{graph}
  %\VignetteKeywords{Graphs}
  %\VignettePackage{RBGL}
  %\VignetteEncoding{UTF-8}
output: 
  BiocStyle::html_document: 
    number_sections: yes 
    toc: yes 
    toc_depth: 4
---

```{r Setup, include = FALSE}
library(RBGL)
library(Rgraphviz)
library(XML)
```

> *Summary*. The `r Biocpkg("RBGL")` package is primarily an interface
> from R to the Boost Graph Library (BGL). It includes some graph
> algorithms built on top of those from BGL and some algorithms
> independent of BGL.

# Basic notations/Preliminaries

## Basics Notations

We use the following notation:

*G*: a graph, represented as G = (V, E); *V* = v1, v2, ..., vn: a set
of vertices (or nodes); *E* = e1, e2, ..., em: a set of edges with ei =
[vj, vk], with vj, vk are in V; *W* = w1, w2, ..., wm: a set of
weights of the edges, i.e., wi is the weight on edge ei.

A *walk* is a sequence of vertices v1, v2, ..., vk such that for all i,
[vi, vi+1] in *E*. A *path* is a walk without repeated vertices. A
*cycle* is a path that begins and ends at the same vertice.

A *directed* graph is a graph with direction assigned to its edges,
therefore, [vj, vk] != [vk, vj].

A *directed acyclic graph (DAG)* is a directed graph with no directed
cycle.

An *in-degree* of vertex v is the total number of edges [u, v] in E;
an *out-degree* of v is the total number of edges [v, u] in *E*.

A network *N* is a directed graph *G* with (a) a source *s* whose
in-degree is 0, (b) a sink *t* whose out-degree is 0, and (c) a
*capacity* for each edge in a network.

A *flow* in *N* assigns a value on each edge that doesn't exceed its
capacity, all the internal vertices have the same incoming flow as the
outgoing flow, *s* has outgoing flow only, *t* has incoming flow only.

## Examples in use

We are going to use the following graphs repeatedly in the examples.

```{r bfDemo}
con <- file(system.file("XML/bfsex.gxl", package="RBGL"))
bf <- fromGXL(con)
close(con)
```

```{r figbf, echo = FALSE, eval = FALSE}
plot(bf, main="a) Breath-First Search Example")
```

```{r dfDemo}
con <- file(system.file("XML/dfsex.gxl", package="RBGL"))
df <- fromGXL(con)
close(con)
```

```{r figdf, echo = FALSE, eval = FALSE}
plot(df, main="b) Depth-First Search Example") 
```

```{r dijkstraDemo}
con <- file(system.file("XML/dijkex.gxl", package="RBGL"))
dijk <- fromGXL(con)
close(con)
```

```{r figdijk, echo = FALSE, eval = FALSE}
plot(dijk, main="c) Dijkstra's Example")
```

```{r connDemo}
con <- file(system.file("XML/conn.gxl", package="RBGL"))
coex <- fromGXL(con)
close(con)
```

```{r figcoex, echo = FALSE, eval = FALSE}
plot(coex, main="d) Coex Example")
```

```{r conn2Demo}
con <- file(system.file("XML/conn2.gxl", package="RBGL"))
coex2 <- fromGXL(con)
close(con)
```

```{r figcoex2, echo = FALSE, eval = FALSE}
plot(coex2, main="e) Coex2 Example")
```

```{r conn2iDemo}
con <- file(system.file("XML/conn2iso.gxl", package="RBGL"))
coex2i <- fromGXL(con)
close(con)
```

```{r figcoex2i, echo = FALSE, eval = FALSE}
plot(coex2i, main="f) Coex2 Isomorphism Example")
```

```{r kmstDemo}
con <- file(system.file("XML/kmstEx.gxl", package="RBGL"))
km <- fromGXL(con)
close(con)
```

```{r figkmst, echo = FALSE, eval = FALSE}
plot(km, main="g) Kruskal MST Example")
```

```{r bicoDemo}
con <- file(system.file("XML/biconn.gxl", package="RBGL"))
bicoex <- fromGXL(con)
close(con)
```

```{r figbico, echo = FALSE, eval = FALSE}
plot(bicoex, main="h) Biconnected Component Example")
```

```{r ospfDemo}
con <- file(system.file("XML/ospf.gxl", package="RBGL"))
ospf <- fromGXL(con)
close(con)
```

```{r figospf, echo = FALSE, eval = FALSE}
plot(ospf, main="i) Ospf Example")
```

```{r zzDemo}
con <- file(system.file("dot/joh.gxl", package="RBGL"))
joh <- fromGXL(con)
close(con)
```

```{r figjoh, echo = FALSE, eval = FALSE}
plot(joh, main="j) joh Example")
```

```{r hcsDemo}
con <- file(system.file("XML/hcs.gxl", package="RBGL"))
hcs <- fromGXL(con)
close(con)
```

```{r fighcs, echo = FALSE, eval = FALSE}
plot(hcs, main="k) HCS Example")
```

```{r kclexDemo}
con <- file(system.file("XML/snacliqueex.gxl", package="RBGL"))
kclex <- fromGXL(con)
close(con)
```

```{r figkclex, echo = FALSE, eval = FALSE}
plot(kclex, main="l) kCliques Example")
```

```{r kcoexDemo}
con <- file(system.file("XML/snacoreex.gxl", package="RBGL"))
kcoex <- fromGXL(con)
close(con)
```

```{r figkcoex, echo = FALSE, eval = FALSE}
plot(kcoex, main="m) kCores Example")
```

```{r layout, echo = FALSE}
layout_matrix_1 <- matrix(1:4, ncol = 2)
```
  
```{r, fig.cap = "The example graphs (I).", fig.height = 8, echo = FALSE}
layout(layout_matrix_1)
<<figbf>>
<<figdijk>>
<<figdf>>
<<figcoex>>
```
  
```{r, fig.cap = "The example graphs (II).", fig.height = 8, echo = FALSE}
layout(layout_matrix_1)
<<figcoex2>>
<<figkmst>>
<<figcoex2i>>
<<figbico>>
```
  
```{r, fig.cap = "The example graphs (III).", fig.height = 8, echo = FALSE}
layout(layout_matrix_1)
<<fighcs>>
<<figkcoex>>
<<figkclex>>
```

<a id="fig4"></a>
```{r fdpic, fig.cap = "File dependency digraph example from Boost", fig.height = 7, results = "hide", echo = FALSE}
<<showFileDep>>
plot(FileDep)
```

# Working with the Bioconductor `graph` class

An example object representing file dependencies is included, as shown
in Figure [4](#fig4).

```{r showFileDep}
data(FileDep)
FileDep
```

```{r, echo = FALSE}
layout_matrix_2 <- matrix(1:2, ncol = 2)
```

```{r figdfsex, fig.cap = "a) The graph for depth-first-search example b) The graph for depth-first-search example, showing search orders.", echo = FALSE, results = "hide"}
layout(layout_matrix_2)
<<DFSdemo>>
<<figdfs1>>
<<figdfs2>>
```

# Algorithms from BGL

## Depth First Search

The `dfs` function returns two vectors of node names of discovery and
finish order in a depth-first-search (DFS), starting at the given
vertex.

```{r DFSdemo}
print(dfs.res <- dfs(df, "y"))
```

In this example, DFS starts with *y*, reaches *x* and *v*; DFS restarts
from *w*, reaches *z*; DFS restarts from *u*; at this point, all the
vertices in the graph are visited once and only once. You could see the
search order in the figure.

```{r figdfs1, echo = FALSE, eval = FALSE}
plot(df, main="a) DFS Example")
```

```{r figdfs2, echo = FALSE, eval = FALSE}
dfsNattrs <- makeNodeAttrs(df) 
dfsNattrs$label[dfs.res$discovered] <- 1:numNodes(df) 
plot(df, nodeAttrs=dfsNattrs, main="b) DFS Example with search order")
```

## Breadth First Search

The `bfs` function returns a vector of node names of discovery order in
a breadth-first search (BFS), starting at the given vertex.

```{r figbfsex, fig.cap = "a) The graph for breadth-first-search example b) The graph for breadth-first-search example, showing search orders.", echo = FALSE, results = "hide"}
layout(layout_matrix_2)
<<BFSdemo>>
<<figbfs1>>
<<figbfs2>>
```

```{r BFSdemo}
print(bfs.res <- bfs(bf,"s"))
```

In this example, BFS starts from vertex *s*, reaches *w*; from *w* BFS
reaches *r*, *t* and *x*; from *r* BFS reaches *v*; from *t* BFS reaches
*u*; from *x* BFS reaches *y*; at this time, BFS visits all the vertices
in the graph once and only once.

```{r figbfs1, echo = FALSE, eval = FALSE}
plot(bf, main="a) BFS Example")
```

```{r figbfs2, echo = FALSE, eval = FALSE}
bfsNattrs <- makeNodeAttrs(bf)
bfsNattrs$label[bfs.res] <- 1:numNodes(bf)
plot(bf, nodeAttrs=bfsNattrs, main="b) BFS Example with search order")
```

## Shortest paths

Edge weights play a major role in shortest-path problems. The weight of
an edge in a graph could represent the relationship between the two
vertices, such as distance, probability, etc.

TO-BE-FINALIZED: Our knowledge of such a relationship between two
vertices is: (1) we know there is an edge and there is a measured value
on it; (2) we know there is an edge but there is NO measured value on
it; (3) we know there is NO edge; (4) we DO NOT know if there is an
edge.

Corresponding edge weights are: case 1: measured value; case 2: `NA`;
case 3: `Inf`; case 4: TO-BE-DETERMINED

When there is a loop of negative weight between two vertices, the
distance between them is `-Inf`.

The shortest path problem is to find a path between two vertices where
the sum of all the edge weights on this path is minimum.

There are two sets of algorithms available: (1) find shortest paths
between a single vertex, say, *source s*, and all other vertices, i.e.,
*V-s*, available algorithms are: *Dijkstra's*, *Bellman Ford's* and
*DAG*, and (2) find shortest paths between all pairs of vertices,
available algorithms are: *Johnson's* and *Floyd Warshall's*.

*Dijkstra's algorithm* is for the single-source shortest-paths problem
on graphs (directed or undirected) with non-negative weights on edges.
If all the edges have the same weight, use depth-first-search instead.

```{r dijkdemo1}
nodes(dijk)
edgeWeights(dijk)
dijkstra.sp(dijk)
```

The function `dijkstra.sp` finds the shortest paths from A, which is the
first node in the node list - default source, to all other vertices in
the graph: *B, C, D, E*, shown in the *distances* part. The *penult*
shows TO-BE-FILLED-IN.

For instance, edge *A->C* exists and carries a weight of 1, so the
shortest path from *A* to *C* is 1; the shortest path from *A* to *B*
goes through *A->C->D->E->B* and has weight of 6 (1+3+1+1).

```{r dijkdemo2}
nodes(ospf)[6]
dijkstra.sp(ospf,nodes(ospf)[6])
sp.between(ospf, "RT6", "RT1")
```

```{r figospfs, fig.width = 6, fig.cap = "Network example from BGL.", echo = FALSE}
plot(ospf)
```

The first part of this example finds the shortest paths from *start RT6*
to all the other vertices in the graph, and the second part finds the
shortest path between two vertices: *RT6* and *RT1*.

*Bellman-Ford's algorithm* is for the single-source shortest-paths
problem on graphs (directed or undirected) with both positive and
negative edge weights. The default source is the first entry in the list
of nodes in the graph.

```{r bellmanfordDemo}
dd <- coex2
nodes(dd)
bellman.ford.sp(dd)
bellman.ford.sp(dd,nodes(dd)[2])
```

The first `bellman.ford.sp` returns the shortest paths from *start A*,
which is the first vertex on the node list, to all other vertices. The
second call shows the shortest paths from *start B* to all other
vertices, since there is no path from *B* to *A*, the *distance* between
them is `Inf`.

The *DAG algorithm* is for the single-source shortest-paths problem on a
weighted, directed acyclic graph (DAG), which is more efficient for DAG
than both Dijkstra's and Bellman-Ford's algorithms. When all the edges
have the same weight, use depth-first-search instead.

```{r DAGDemo}
dd <- coex2
dag.sp(dd)
dag.sp(dd,nodes(dd)[2])
```

It's easy to see that *conn2.gxl* doesn't contain any cycle, so we could
use function `dag.sp` on it. The first example finds the shortest paths
from the *start A* to all other vertices. The second example finds the
shortest paths from *start B* to all other vertices, since no path goes
from *B* to *A*, the distance is *Inf*.

*Johnson's algorithm* finds the shortest path between every pair of
vertices in a sparse graph. Its time complexity is *O(V E log V)*.

```{r johnsonDemo}
zz <- joh
edgeWeights(zz)
johnson.all.pairs.sp(zz)
```

This example uses a graph with negative edge weights.

The shortest paths between all pairs of vertices are presented in the
matrix, entry [i, j] gives the distance from vertex *i* to vertex *j*.
For example, the shortest path from vertex *c* to vertex *d* is of
length 5; the shortest path from vertex *a* to vertex *e* is of length
-4, since edge *a->e* is available and of distance -4; the shortest
distance from *a* to *c* is -3.

*Floyd-Warshall's algorithm* finds the shortest path between every pair
of vertices in a dense graph.

```{r floydwarshallDemo}
floyd.warshall.all.pairs.sp(coex)
```

All edge distances are assumed to be 1, if not given. Since the graph is
undirected, the distance matrix is symmetric, for example, distance from
*C* to *G* is the same as that from *G* to *C*.

```{r figjohn, fig.cap = "Example: Johnson-all-pairs-shortest-paths example", fig.small = TRUE, echo = FALSE} 
plot(zz)
```

## Minimum spanning tree

Minimum-spanning-tree (MST) problem is to find a subset of edges that
connect all the vertices, contains no cycles and have the minimum weight
sum.

There are two algorithms available: *Kruskal's algorithm* and *Prim's
algorithm*. Both are for undirected graphs with weighted edges, and both
return a list of edges, weights and nodes determining MST.

The `mstree.kruskal` function finds the MST by Kruskal's algorithm.

```{r KMSTdemo}
mstree.kruskal(km)
```

This graph is treated as undirected graph with corresponding weights.
MST consists of 4 edges, *A->C, D->E, E->A, B->D*, each is of weight 1.

The `mstree.prim` function finds the MST by Prim's algorithm.

```{r primDemo}
mstree.prim(coex2)
```

The graph is treated as undirected graph with default weight 1. MST
consists of 7 edges, *A->B, A->C, A->D, C->E, D->H, E->F, C->G*.

```{r figkm, fig.cap = "Kruskal MST examples.", echo = FALSE, results = "hide"}
<<conndemo>>
layout(layout_matrix_2)
<<figkmst>>
<<figkm1>>
```

## Connected components 

There are several algorithms available for this group of problems.

A *connected component* of an undirected graph is a subgraph that for
any two vertices in this subgraph, *u* and *v*, there's a path from *u*
to *v*.

The `connectedComp` function computes the connected components in an
undirected graph.

```{r conndemo} 
km1 <- km 
km1 <- graph::addNode(c("F","G","H"), km1) 
km1 <- addEdge("G", "H", km1, 1) 
km1 <- addEdge("H", "G", km1, 1) 
connectedComp(ugraph(km1))
```

```{r figkm1, echo = FALSE, eval = FALSE}
plot(km1, main="Modified Kruskal MST example")
```

The original graph has one connected component. After we add three
vertices, *F, G, H* and an edge *G-H*, make the graph *undirected*, the
modified graph has three connected components now.

A *strongly connected component* of a directed graph is a connected
subgraph that for every pair of vertices in this subgraph, *u* and *v*,
there are both a path from *u* to *v* and a path from *v* to *u*.

The `strongComp` function computes the strongly connected components in
a directed graph.

```{r sconndemo}
km2 <- km
km2 <- graph::addNode(c("F","G","H"), km2)
km2 <- addEdge("G", "H", km2, 1)
km2 <- addEdge("H", "G", km2, 1)
strongComp(km2)
```

After adding three vertices, *F, G, H* and an edge *G-H*, there are
three strong components in the graph now.

A *biconnected* graph is a connected graph that removal of any single
vertex doesn't disconnect it. If the removal of a vertex increases the
number of components in a graph, this vertex is call an *articulation
point*.

The `biConnComp` function computes the biconnected components in an
undirected graph. The `articulationPoints` function finds all the
articulation points in an undirected graph.

```{r biConnCompdemo} 
biConnComp(bicoex) 
articulationPoints(bicoex)
```

```{r figbicoex, fig.width = 6, fig.cap = "Biconnected components example from Boost library.", echo = FALSE}
plot(bicoex)
```

There are 4 biconnected components in the example: one with vertices *B, C, D* 
and edges *B-C, C-D, B-D* labeled 0, one with vertices *A, B, E, F* 
and edges *A-B, B-E, E-F, A-F* labeled 1, one with vertices *G, H, I*
and edges *G-I, G-H, I-H* labeled 2, and one with vertices *A, G* and
edges *A-G* labeled 3.

There are 3 articulation points in the example: *A, B, G*. It's easy to
see removing any one of them will result in more connected components.

When you *add* edges to an undirected graph and want to get updated
information on the connected components, you could use the following
functions: `init.incremental.components` function to initialize the
process; after adding edges to the graph, use `incremental.components`
function to update the information on the connected components; use
`same.component` function to find out if two vertices are in the same
connected component.

Currently, only one incremental graph is allowed at any given time. To
start on a new graph, you need to call `init.incremental.components`
first.

```{r incrCompdemo} 
jcoex <- join(coex, hcs)
x <- init.incremental.components(jcoex)
incremental.components(jcoex)
same.component(jcoex, "A", "F")
same.component(jcoex, "A", "A1")
jcoex <- addEdge("A", "A1", jcoex) 
x <- init.incremental.components(jcoex)
incremental.components(jcoex)
same.component(jcoex, "A", "A1")
```

```{r figjcoex, fig.width = 6, fig.cap = "Example on incremental components: a graph connecting coexand hcs.", echo = FALSE}
plot(jcoex)
```

In the first part of this example, we join two separate graphs together,
the resulting graph contains two connected components. Vertices *A* and
*F* are in the same connected component, while vertices *A* and *A1* are
not in the same connected component.

In the second part of the example, we add an edge connecting *A* and
*X*, which effectively connects the two subgraphs, we have only one
connected component left, which consists of all the vertices from the
two original graphs, *A* and *A1* are in the same connected component
now.

## Maximum Flow

The functions, `edmonds.karp.max.flow` and `push.relabel.max.flow` are
available to find the maximum flow between source and sink.

```{r MaxFlowdemo}
edgeWeights(dijk) 
edmonds.karp.max.flow(dijk, "B", "D") 
push.relabel.max.flow(dijk, "C", "B")
```

Call to `edmonds.karp.max.flow` finds the maximum flow of 2 from *B* to
*D*: one part of flow 1 is *B -> D* directly, another part of flow 1 is
*B -> E -> A -> C -> D*.

Call to `push.relabel.max.flow` find the maximum flow of 8 from *C* to
*B*: one part of flow 7 is *C -> B* directly, another part of flow 1 is
*C -> D -> E -> B*.

You can see the flow on each edge in the output, and each is no more
than the capacity of the edge.

## Sparse Matrix Ordering

There are three functions available in this category:
`cuthill.mckee.ordering`, `minDegreeOrdering` and `sloan.ordering`.

*Cuthill-McKee's algorithm* tries to reduce the bandwidth of a graph by
renumbering its vertices. The outputs are the vertices in the new
ordering and reverse ordering.

*Minimum degree Ordering* is one approach that tries to reduce fill-ins
in matrix reordering, which turns a system of equations *A x = b* to 
*(P A PT)(P x) = P b*.

*Sloan Ordering* tries to reduce the profile and wavefront of a graph by
renumbering its vertices.

```{r SparseMatrixOrderingdemo}
dijk1 <- ugraph(dijk)
cuthill.mckee.ordering(dijk1) 
minDegreeOrdering(dijk1)
sloan.ordering(dijk1)
```

TODO: EXPLAIN THESE OUTPUT.

## Edge connectivity and minimum disconnecting set

For a single connected undirected graph, function *edgeConnectivity*
calculates the minimum number of edges that have to be removed to create
two disconnected components. No edge weight is taken into account and
the output is the edges that need to be removed.

This is very similar to the *minCut* algorithm, which takes the edge
weights into account when removing edges and outputs the vertices on the
two disconnected components.

```{r edgeConndemo} 
edgeConnectivity(coex)
```

Mimimum of two edges must be removed to create two disconnected
components: edges *D-E* and *D-H*.

## Topological sort

The `tsort` function will return the names of vertices from a DAG in
topological sort order.

```{r tsortDemo1}
tsort(FileDep)
```

Note that if the input graph is not a DAG, BGL `topological_sort` will
check this and throw 'not a dag'. This is crudely captured in the
interface (a message is written to the console and zeroes are returned).

```{r tsortDemo2} 
FD2 <- FileDep 
# now introduce a cycle 
FD2 <- addEdge(from="bar_o", to="dax_h", FD2) 
tsort(FD2)
```

## Isomorphism

The `isomorphism` function determines if two graphs are isomorphism,
i.e., determines if there is a one-to-one mapping *f* of vertices from
one graph *g1* to the other *g2* such that edge *u -> v* is in *E(g1)*
if edge *f(u) -> f(v)* exists and is in *E(g2)*.

```{r Isomorphismdemo}
isomorphism(dijk, coex2) 
isomorphism(coex2i, coex2) 
```

The function handles both directed and undirected graphs. There are more
vertices in graph *conn2* than *dijkstra*, so it's impossible to find a
one-to-one mapping between them. One the other hand, graph *conn2i* is
basically the same graph as *conn2* except the vertices have different
names, so they are isomorphism.

## Vertex Coloring

The `sequential.vertex.coloring` function assigns colors, as numbers 0,
1, 2, ..., to vertices in a graph so that two vertices connected by an
edge are of different colors. It does not guarantee that minimum number
of colors is used, and the result depends on the input ordering of the
vertices in the graph.

```{r VertexColoringdemo}
sequential.vertex.coloring(coex)
```

We need 4 colors for the vertices of this graph, one color scheme is to
give color 0 to vertices *A, E*, color 1 to vertices *B, H*, color 2 to
vertices *C, F* and color 3 to vertices *D, G*.

```{r, fig.small = TRUE, fig.cap = "Example conn2i", echo = FALSE}
plot(coex2i)
```

## Wavefront, Profiles

TODO: EXPLAIN THESE TERMS

The following functions are available: `ith.wavefront`, `maxWavefront`,
`aver.wavefront` and `rms.wavefront`.

```{r wavefrontdemo} 
ss <- 1 
ith.wavefront(dijk, ss)
maxWavefront(dijk) 
aver.wavefront(dijk) 
rms.wavefront(dijk)
```

TODO: EXPLAIN THESE RESULTS

## Betweenness Centrality and Clustering

*Betweenness centrality* of a vertex (or an edge) measures its
importance in a graph, i.e., among all the shortest paths between every
pair of vertices in the graph, how many of them have to go through this
vertex (or edge). *Relative* betweenness centrality is calculated by
scaling the *absolute* betweenness centrality by factor
*2/((n-1)(n-2))*, where *n* is the number of vertices in the graph.

The `brandes.betweenness.centrality` function implements Brandes'
algorithm in calculating betweenness centrality.

The `betweenness.centrality.clustering` function implements clustering
in a graph based on edge betweenness centrality.

TODO: EXPLAIN MORE

```{r Centralitydemo, echo = TRUE} 
brandes.betweenness.centrality(coex)
betweenness.centrality.clustering(coex, 0.1, TRUE)
```

# Algorithms built on RBGL

## Min-Cut

Given an undirected graph G=(V, E) of a single connected component, a
*cut* is a partition of the set of vertices into two non-empty subsets S
and V-S, a *cost* is the weight sum of edges that are incident on one
vertex in S and one vertex in V-S. The min-cut problem is to find a cut
(S, V-S) of minimum cost.

For simplicity, subset *S* is the smaller of the two.

```{r mincutdemo} 
minCut(coex)
```

Currently all edge weights are assumed to be 1, minimum cut is of weight
2, it will partition the graph into two subsets: subset *A, B, C, D* and
subset *E, H, F, G*.

## highlyConnSG

A graph *G* with *n* vertices is highly connected if its connectivity
*k(G) > n/2*. Function *highlyConnSG* partitions a graph into a set of
highly connected subgraphs, by using minimum-cut algorithm repeatedly.
To improve performance, it takes special care of singletons, low degree
vertices and merges clusters.

```{r highlyConnSGdemo}
highlyConnSG(coex) 
highlyConnSG(hcs)
```

In graph *conn*, two highly-connected-subgraphs are found: subgraph with
vertices *A, B, C, D* and subgraph with vertices *E, H, F, G*.

In graph *hcs*, 3 highly-connected-subgraphs are found: subgraph with
vertices *A1, A2, A3, A4, A5*, subgraph with vertices *B1, B2, B3, B4*
and subgraph with vertices *X, Y, Z*.

# Algorithms independent from RBGL

## maxClique

A *clique* is a complete subgraph, i.e., there is an edge between every
pair of vertices.

*Maximum Clique* problem is to find the largest clique in a graph. This
problem is NP-complete, which means it cannot be solved by any known
polynomial algorithm.

Function *maxClique* implements the algorithm from *Finding all cliques
of an undirected graph*, by C. Bron and J. Kerbosch (CACM, Sept 1973,
Vol 16, No. 9.), which finds all the cliques in a graph.

```{r MaxCliquedemo} 
maxClique(coex) 
maxClique(hcs)
```

In graph *conn*, 3 cliques are found: clique with vertices *D, B, C, A*,
clique with vertices *D, E, H* and clique with vertices *F, E, H, H*.

In graph *hcs*, 10 cliques are found. For instance, vertices *A2, A4,
A3* form a clique, vertices *B1, Y* form a clique.

## is.triangulated

A graph is *triangulated* if all cycles of length 4 or more have a
chord. The `is.triangulated` function returns TRUE or FALSE,
accordingly.

We implemented the following algorithm from *Combinatorial Optimization:
algorithms and complexity* (p. 403) by C. H. Papadimitriou, K.
Steiglitz: G is chordal if either G is an empty graph, or there is a
*v* in *V* such that (i) the neighborhood of *v*, i.e., *v* and its
adjacent vertices, forms a clique, and (ii) recursively, *G-v* is chordal.

```{r IsTriangulateddemo}
is.triangulated(coex)
is.triangulated(hcs)
```

## separates

Function *separates* determines if a subset of vertices separates two
other subsets of vertices, and returns TRUE or FALSE, accordingly.

```{r Separatesdemo} 
separates("B", "A", "E", km)
separates("B", "A", "C", km)
```

## kCores

A *k-core* in a graph is a subgraph where each vertex is adjacent to at
least `k` other vertices in the same subgraph.

Function *kCores* finds all the k-cores in a graph. It returns the core
numbers for all the nodes in the graph. When the given graph is
directed, you can choose whether in-degree, out-degree or both should be
considered.

The k-core of a graph is not a necessarily connected subgraph. If i > j, 
the i-core of a graph contains the j-core of the same graph.

The implementation is based on the algorithm by V. Batagelj and M.
Zaversnik, 2002.

```{r kCoresdemo1} 
kCores(kcoex) 
kcoex2 <- coex2 
kCores(kcoex2)
kCores(kcoex2, "in") 
kCores(kcoex2, "out") 
g1 <- addEdge("C", "B", kcoex2) 
kCores(g1, "in") 
g2 <- addEdge("C", "A", kcoex2)
kCores(g2, "out")
```

```{r figkcores, fig.small = TRUE, fig.cap = "K-cores Example.", echo = FALSE}
plot(kcoex)
```

The example on directed graph, "conn2", turns out to be a
waterfall-like graph. If we order the nodes as: A, B, C, D, E, F, H, G,
all the edges go in the same direction, i.e., i -> j, i < j.

Let's consider in-degree-only case: A has no in-edge so it is 0-core;
after you eliminate A, no in-edge to B, so B is 0-core; continue this,
we could see that there's no subset of nodes that each and every single
node has 1 in-degree. Therefore, they are all of 0-core.

For out-degree-only case: G has no out-edge, so it's 0-core; after
eliminating G, F has no out-edge, so F is 0-core; continue this process,
we could see that there's no subset of nodes that each and every single
node has 1 out-edge. Therefore, they are all of 0-core.

If we add edge(s) to break the waterfall-like property, *C->B*,
*C->A*, separately, we could see the changes in the core numbers that
are consistant with the analysis above.

## kCliques

## kCliques

In social network analysis, a k-cliques is a maximum subgraph that the
shortest distance between any two nodes is no more than k.

Function *kCliques* finds all the k-cliques in an undirected graph (k =
1, ..., N, where N is the length of the longest shortest-path). It
returns all the k-cliques.

Let D be a matrix, D[i][j] is the shortest path from node i to node
j. Algorithm is outlined as following. o. use Johnson's algorithm to
fill D; let N = max(D[i][j]) for all i, j; o. each edge is a
1-clique by itself; o. for k = 2, ..., N, try to expand each
(k-1)-clique to k-clique: o. consider a (k-1)-clique the current
k-clique KC; o. repeat the following: if for all nodes j in KC,
D[v][j] <= k, add node v to KC; o. eliminate duplicates; o. the
whole graph is N-clique.

```{r kCliquesdemo} 
kCliques(kclex)
```

```{r figkcliques, fig.small = TRUE, fig.cap = "K-cliques Example", echo = FALSE}
plot(kclex)
```